Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2017-10-24 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex
typos

CheriBSD, an operating sytem built on the CHERI capability machine, uses per-component separate stacks and a central, trusted stack manager component to enforce well-bracketed control flow and private state encapsulation \citep{watson_cheri:_2015}.
To prevent the stack pointer (or derivatives from it) from being passed to other components, CheriBSD uses local capabilities: a type of capabilities that can be kept in registers but not stored in memory (except through store-local capabilities like the stack pointer itself).
The literature does not contain many details about the mechanism, let alone a formal analysis, but most importantly, the mechanism does not seem like it can support higher-order interfaces (e.g. C function pointers) in a scalable fashion.

@@

\section{Linear Capabilities, not Local}
What we essentially want to achieve in a secure calling convention is to give an invoked function access to a stack pointer and a return pointer, but only temporarily: during the execution of the invoked function. 
In our original approach, we make both pointers local, so that the CHERI processor guarantees, essentially, that they can only be stored in registers or on the stack.
To make sure that an invoked function has no way of transferring the pointers accross outside the lifetime of its stack frame, it then suffices to block only these two communication channels, rather than the entire memory.
In other words, our approach relies on clearing registers and the stack before we pass control to untrusted code, as a way of preventing it from obtaining access to other stack and return pointers than the ones for the stack frame it is invoked in.
The requirement for clearing registers is unproblematic, but clearing the entire stack upon every boundary crossing is a non-trivial requirement.
We speculate that it can be made efficient with custom hardware extensions\footnote{Specifically, we imagine a kind of pre-L0 processor cache that does not retain values but only remembers that a certain range has been zeroed out.}, but this is not obvious.

>>>>

CheriBSD, an operating system built on the CHERI capability machine, uses per-component separate stacks and a central, trusted stack manager component to enforce well-bracketed control flow and private state encapsulation \citep{watson_cheri:_2015}.
To prevent the stack pointer (or derivatives from it) from being passed to other components, CheriBSD uses local capabilities: a type of capabilities that can be kept in registers but not stored in memory (except through store-local capabilities like the stack pointer itself).
The literature does not contain many details about the mechanism, let alone a formal analysis, but most importantly, the mechanism does not seem like it can support higher-order interfaces (e.g. C function pointers) in a scalable fashion.

@@

\section{Linear Capabilities, not Local}
What we essentially want to achieve in a secure calling convention is to give an invoked function access to a stack pointer and a return pointer, but only temporarily: during the execution of the invoked function. 
In our original approach, we make both pointers local, so that the CHERI processor guarantees, essentially, that they can only be stored in registers or on the stack.
To make sure that an invoked function has no way of transferring the pointers across outside the lifetime of its stack frame, it then suffices to block only these two communication channels, rather than the entire memory.
In other words, our approach relies on clearing registers and the stack before we pass control to untrusted code, as a way of preventing it from obtaining access to other stack and return pointers than the ones for the stack frame it is invoked in.
The requirement for clearing registers is unproblematic, but clearing the entire stack upon every boundary crossing is a non-trivial requirement.
We speculate that it can be made efficient with custom hardware extensions\footnote{Specifically, we imagine a kind of pre-L0 processor cache that does not retain values but only remembers that a certain range has been zeroed out.}, but this is not obvious.
